Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-10 fib2.tex
quite a few miscellaneous edits. its now a few paragraphs too long again.

Topologically ordered quantum systems in two dimensions show tremendous promise for 
long-term storage and processing of quantum information~\cite{Kitaev2003, Dennis2002, Nayak2008}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and have quasiparticle excitations 
exhibiting anyonic statistics~\cite{Wilczek1990}. These systems can in general be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

Quantum error correction is vital to harnessing the computational power of topologically 
ordered systems. When coupled to a heat bath at any finite temperature, thermal fluctuations 
will create spurious anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
at low temperature must be augmented by an \emph{active} decoding procedure.

@@

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015b}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction 
in these systems. When considering more general anyon models, their 
ability to perform universal quantum computation would seem a significant 
barrier to their simulation on a classical computer. While simulation 
of general dynamics does indeed seem intractable, we argue that 
the kinds of processes that are typical of thermal noise 
are sufficiently structured  to allow for their classical simulation 
in the regimes where we expect successful error correction to 
be possible. This insight allows us to simulate the noise 
and recovery processes for a quantum code based on a universal anyon model.

@@

We consider a 
2-dimensional system with
point-like excitations or quasiparticles known as anyons.
These anyons show exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by \dude{directed} closed loops (Wilson loops).
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. % (UMTC). 
However, from the perspective of
\emph{topological quantum field theory}
anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid,
but the manifold in which the anyons reside 
that becomes deformed (twisted) by these braid moves.
This is the perspective we adopt (see \Fref{f:braidloop}).
\dude{Fig or FIG ?}
%In particular, this shows the action of braiding
%on arbitrary observables (Wilson loops) as
%these deform with the manifold.

@@

\begin{figure}[t!]
\begin{center}
    \includegraphics[]{pic-braid-loop.pdf}
\caption{
%The perspective of TQFT.
Here we show the world lines of a braid.
%This braid relates two different ways of observing the same charges.
In this case, we see how it relates two different ways of observing (or fusing) the top two charges.
%This braid relates two different ways of observing the same charges.
%We are free to draw these loops around charges,
In general braiding will twist these loops in different ways
according to how the manifold itself is being deformed.
\label{f:braidloop}
}
\end{center}
\vspace{-10pt}
\end{figure}

%The state \dude{space?} of a collection of anyons
%can be described by the observables of the system.
The \dude{key?} difference between abelian and non-abelian
anyon theories is that the outcomes of measurements in an abelian theory
are determined uniquely particle content only.
%The defining characteristic of a \emph{non-abelian}
%anyon theory is that ...
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for non-abelian anyons.
In particular, we consider a system supporting (non-abelian) Fibonacci
anyon excitations, denoted by $\tau.$
Two such anyons can have total charge
that is a superposition of $\tau$ and $\vac$ (vacuum),
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio.

@@

A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$.

@@

@@

For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\vac}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\vac,\tau)$.

Finally, there is a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
This global space is the one in which we will encode, and demonstrate protection of, quantum 
information. 
For detailed description of these 
processes in Fibonacci anyons, see e.g.~Ref.~\cite{Nayak2008} and references therein. 
Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be 
sufficient to implement universal quantum computation~\cite{Freedman2002, Nayak2008}.

This treatment of anyon dynamics we consider to be a phenomenological model;
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
\dude{but how do we relate to discrete models: toric code, string nets, etc.
microscopic lattice models.}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
%We tile the manifols with  a finite set of $L\times L$ Wilson loops arranged in
These \emph{tiles} will be the observables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{T\in \Lambda}\proj{\vac}_T\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_T$ the projector to charge $\vac$ at tile $T$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space contains the two-fold degenerate codespace.
%We identify this ground space as the codespace of our model.

Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when 
there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure in the noise 
processes that allows us to classically simulate them.

@@

Although simulating pair creation, braiding, and fusion of $n$ Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure that we can 
exploit to efficiently simulate typical processes of interest. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons 
(we leave a more rigorous analysis of simulability for noise 
and error correction processes as an open problem).

@@

However, random noise processes are not the only dynamics that we need to consider. We 
must also consider the effect of the error-correction routine itself. This acts iteratively to fuse 
anyons on increasing length scales. While this kind of fusion would typically merge components, 
forcing us to compute dynamics of larger and larger sets of anyons,
large components are sparsely distributed
(and thus unlikely to be merged), and in addition at each length scale the total number of 
anyons present is dramatically reduced by fusion, leading to a smaller number of anyons that 
must be simulated.

In the regime where the action of noise plus decoder does not percolate
it is clear \dude{check this} that the simulation is efficient.
However, with strong enough noise the state will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics %for finite lattice size, we use heuristics
that minimize non-local braid moves, which helps to
maintain the state as a separable product beyond the
sub-percolated regime.
This enables simulation of error correction in
the area around decoder threshold for lattice sizes up to $L=128.$

@@

In order to keep track of the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra piece of curve to the lattice supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$ and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
%\dude{Heisenberg picture?}
%The observable corresponding to the tile is then
%represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
%\cggb{This paragraph and the next still needs another look.}

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}
\caption{
(a) Noise processes initially form isolated sets of pair-created anyons, 
each crossing the boundary of a tile. 
(b) To measure the total charge 
contained within each tile, 
we first join the participating curve 
diagrams arbitrarily into a single curve diagram.
(c) The total charge of each tile can then be found 
by braiding anyons around each other until all charges within 
a tile are neighbors on the curve, as in (d).  
The red lines correspond to the worldline braids that perform this step.
}
\label{f:syndrome}
\end{center}
\vspace{-10pt}
\end{figure}

Following the result of these charge measurements, the decoder determines 
a recovery operation that involves fusion of subsets of anyons. 
These fusion results can be calculated in the same way, 
and the output charge placed at an appropriate point in the lattice. 
This procedure is iterated until either the decoder terminates successfully, 
or the simulation itself declares failure.

@@

We plot the performance of the decoder as a function of error rate for varying lattice sizes in 
\Fref{f:threshold}. 
The error rate is parameterized by the Poisson process duration $t_{\mathrm{sim}}$, representing the expected number of errors per edge during the simulation. 
We find evidence of a decoding threshold below which decoding succeeds with asymptotic 
certainty as the system size increases at $t_{\mathrm{sim}}\simeq 0.125 \pm 0.003$.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons accross the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated by an optimal decoder. 
However, calculations of such nontrivial operations suggest that all such 
processes will indeed result in a logical or leakage error.
As such it is likely that neglecting full simulation for 
such events does not significantly affect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon model. 
Though we have chosen several properties of our model and 
simulation in a convenient way for simplicity,
Ref.~\cite{Brell2013} presents good evidence that it is
unlikely that these choices will affect our results qualitatively.
%it seems highly
%unlikely that they will affect the qualitative results, similarly to the results of Ref.~\cite{Brell2013}. 
In particular, although we have modeled our logical qubit as 
encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred anyons. 
This situation would be appropriate to model error-correction routines for topological quantum computation. 
Additionally, we expect our results to be stable to changes in details of the noise model and decoding algorithm.

None of our techniques are restricted to simulation of Fibonacci anyon dynamics, and could 
equally well be used to simulate successful error-correction protocols in an arbitrary anyon code. 
As such, our methods could be used to demonstrate successful 
error-correction in an arbitrary anyonic topological quantum computation.

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, 
applying these methods to more realistic models such as concrete 
microscopic spin models or models with non-topological features would give 
direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. 
It would also be interesting to find \emph{classical} spin models 
whose phase diagram encodes the threshold for error correction in these systems~\cite{Dennis2002}.
%Another potential direction is to incorporate tensor network methods for 
%simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
%Finally, perhaps the most important open question is whether 
Finally, we could ask whether 
\emph{fault-tolerant} decoders can be designed for nonabelian anyon codes.

>>>>

Topologically ordered quantum systems in two dimensions show tremendous promise for 
long-term storage and processing of quantum information~\cite{Kitaev2003, Dennis2002, Nayak2008}. 
The topological features of such systems are insensitive to local 
perturbations~\cite{Bravyi2010, Bravyi2011a, Michalakis2013}, and they have quasiparticle excitations 
exhibiting anyonic statistics~\cite{Wilczek1990}. These systems can in general be used as 
quantum memories~\cite{Kitaev2003, Dennis2002} or to perform universal topological 
quantum computation~\cite{Freedman2002, Nayak2008}.

Quantum error correction is vital to harnessing the computational power of topologically 
ordered systems. When coupled to a heat bath at any non-zero temperature, thermal fluctuations 
will create spurious anyons that diffuse and quickly corrupt the stored quantum 
information~\cite{Pastawski2010}. Thus, the passive protection provided by the mass gap 
at low temperature must be augmented by an \emph{active} decoding procedure.

@@

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015b}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal for quantum computation. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction 
in these systems. When considering more general anyon models, their 
ability to perform universal quantum computation would seem a significant 
barrier to their simulation on a classical computer. While simulation 
of general dynamics does indeed seem intractable, we argue that 
the kinds of processes that are typical of thermal noise 
are sufficiently structured  to allow for their classical simulation 
in the regimes where we expect successful error correction to 
be possible. This insight allows us to simulate the noise 
and recovery processes for a quantum code based on a universal anyon model.

@@

We consider a 
2-dimensional system with
point-like anyonic quasiparticle excitations, which exhibit exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on (the Hilbert space of) degrees of freedom known as the \emph{fusion space.}
To access this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by \dude{directed} closed loops (Wilson loops).
The possible results of such a charge measurement correspond to different anyon particle types in the model.
%The algebraic properties of these processes
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. % (UMTC). 
However, from the perspective of
\emph{topological quantum field theory}
anyons are modelled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid,
but the manifold in which the anyons reside 
that becomes deformed (twisted) by these braid moves.
This is the perspective we adopt (see \Fref{f:braidloop}).
\dude{Fig or FIG ?}
%In particular, this shows the action of braiding
%on arbitrary observables (Wilson loops) as
%these deform with the manifold.

@@

\begin{figure}[t!]
\begin{center}
    \includegraphics[]{pic-braid-loop.pdf}
\caption{
%The perspective of TQFT.
The deformation of observables corresponding to the implementation of a braid process on three anyons is shown. The configurations before and after are connected by world lines shown in red, while the observable Wilson loops are shown in black.
%Here we show the world lines of a braid.
%This braid relates two different ways of observing the same charges.
%In this case, we see how it relates two different ways of observing (or fusing) the top two charges.
This process relates two different ways of observing (or fusing) the top two charges.
%This braid relates two different ways of observing the same charges.
%We are free to draw these loops around charges,
In general, braiding will wind or unwind the Wilson loops, following
the deformation of the manifold itself.
\label{f:braidloop}
}
\end{center}
\vspace{-10pt}
\end{figure}

%The state \dude{space?} of a collection of anyons
%can be described by the observables of the system.
The \dude{key?} difference between abelian and nonabelian
anyon theories is that outcomes of total charge measurements in an abelian theory
are determined uniquely by particle content only.
In contrast, outcomes for nonabelian charge measurements depend on the history of the particles as well as their type.
%The defining characteristic of a \emph{nonabelian}
%anyon theory is that ...
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for nonabelian anyons.
In particular, we consider a system supporting (nonabelian) Fibonacci
anyon excitations, denoted by $\tau$.
Two such anyons can have total charge
that is either $\tau$ or $\vac$ (vacuum), or any superposition of these, and
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$ for $\varphi=\frac{1+\sqrt{5}}{2}$ 
the golden ratio.

@@

@@

A clockwise braiding process is represented by an $R$-move:
\begin{align}
\label{e:rmove}
\includegraphics[]{pic-rmove.pdf}
\end{align}
This is meant to show a \emph{half-twist} of the space
inside the observable with value $c$
holding two anyons (punctures) with charges $a$ and $b$. \cggb{This statement isn't quite true. I suggest just deleting.}

@@

For the Fibonacci anyons, the non-trivial $R$ and $F$ moves are 
\begin{equation*}
	R_{\vac}^{\tau\tau} = \e^{\frac{-4\pi i}{5}} 
	\ \ , \ \
	R_\tau^{\tau\tau}= \e^{\frac{3\pi i}{5}} 
	\ \ , \ \
	F_{\tau}^{\tau\tau\tau} = \begin{pmatrix}\varphi^{-1}&\varphi^{-\frac{1}{2}}\\\varphi^{-\frac{1}{2}}&-\varphi^{-1}\end{pmatrix} \,,
\end{equation*}
where the matrix is given in a basis labelled $(\vac,\tau)$. For more details of the Fibonacci anyon theory, see e.g.~Ref.~\cite{Nayak2008} and references therein. 
Braiding of Fibonacci anyons and measurements of fusion outcomes is known to be 
sufficient to implement universal quantum computation~\cite{Freedman2002, Nayak2008}.

Finally, in addition to the fusion space of any anyons present, there is also a global degeneracy associated with the topology of the 
manifold on which the anyons reside. We will consider systems with the topology of a torus, 
which for the Fibonacci anyons gives rise to a 2-fold degeneracy.
This global space is the one in which we will encode, and demonstrate protection of, quantum 
information.

We consider this treatment of anyon dynamics to be a phenomenological model in that
it neglects any microscopic details of the system.
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Additionally, while we model a continuum anyon theory, our analysis can be trivially reinterpreted to apply directly to lattice spin models or other systems where space is fundamentally discrete.
%\dude{but how do we relate to discrete models: toric code, string nets, etc.
%microscopic lattice models.}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system.
We endow this manifold with a finite set of $L\times L$ Wilson loops arranged in
a square lattice tiling $\Lambda$.
%We tile the manifols with  a finite set of $L\times L$ Wilson loops arranged in
These \emph{tiles} will be the observables accesible to the error-correction procedure (\emph{decoder}).
We can use these observables to construct an idealized Hamiltonian for this system of the form
\begin{align}
	H=-\sum_{t\in \Lambda}\proj{\vac}_t\;,\label{e:hamiltonian}
\end{align}
with $\proj{\vac}_t$ the projector to charge $\vac$ at tile $t$, i.e.~the ground 
space of the model has vacuum total charge in each tile.
Given the toroidal boundary 
conditions, this space \cggb{is} the two-fold degenerate codespace. \cggb{Subtlety in that there could be anyons around in the ground space.}
%We identify this ground space as the codespace of our model.

Typical noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons.
It was seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study.

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015b},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the charge on 
each tile of the lattice, producing a \emph{syndrome} 
of occupied sites. Following this, it forms clusters of nearby occupied tiles and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when 
there is at most a single cluster remaining (see \Fref{f:decode}).

Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query the system by measuring these charges.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or 
when it performs a homologically nontrivial operation, thereby altering the 
codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that 
our qualitative results may be reproduced by most alternative families 
of decoders, although it may not be clear in all cases how to extend these decoders to the nonabelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering 
scheme is compatible with the structure in the noise 
processes that allows us to classically simulate them.

@@

Although simulating pair creation, braiding, and fusion of Fibonacci anyons is equivalent 
in computational power to universal quantum computing (and thus unlikely to be classically 
tractable), noise processes and error-correction procedures have structure that we can 
exploit to efficiently simulate typical processes of interest. In particular, those 
processes in which we expect error-correction to succeed are also those that we expect to 
be able to efficiently simulate for the following heuristic reasons 
(we leave a more rigorous analysis of simulability for noise 
and error correction processes as an open problem).

@@

However, random noise processes are not the only dynamics that we need to consider. We 
must also consider the effect of the error-correction routine itself. This acts iteratively to fuse 
anyons on increasing length scales. While this kind of fusion would typically merge components, 
forcing us to compute dynamics of larger and larger sets of anyons,
large components are sparsely distributed
(and thus unlikely to be merged), and in addition at each length scale the total number of 
anyons present is dramatically reduced by fusion, leading to a smaller number of anyons that 
must be simulated. \cggb{is this really helping the simulation so much as it seems? is it clear that we don't still need to simulate the entire set of anyons even though they arent there anymore?}

In the regime where the combined action of noise and error-correction does not percolate
it is clear \dude{check this} that the simulation is efficient.
However, with strong enough noise the state in the fusion space will
no longer decompose at all and computing dynamics will
become exponentially difficult in the system size.
Despite this, we use heuristics %for finite lattice size, we use heuristics
that minimize non-local braid moves \cggb{what does minimize mean here?}, which helps to
maintain the state as a separable product beyond the
sub-percolated regime.
This enables simulation of error correction in
the regime around the error-correction threshold for linear lattice sizes up to $L=128$.

@@

In order to track the state of the system we maintain 
a set of disjoint directed curves, 
one for each connected component with trivial total charge.
The application of a pair-creation noise process corresponds to the 
addition of an extra curve to the lattice, supporting
two new anyons.
%along with two preferred points on it denoting the anyon locations. 
Following the noise processes, our simulation must %proceed to 
measure the total charge within each tile; the results of 
these measurements will form the error syndrome. 
This requires joining (arbitrarily) any curves that intersect that tile, 
and then performing $R$- and $F$-moves on the resulting
curve so that 
all anyons within the tile have been 
localized within a contiguous region of the curve. 
%\dude{Heisenberg picture?}
%The observable corresponding to the tile is then
%represented in the state \dude{polish}.
%The total charge is then (probabilistically) read off the 
%relevant section of the fusion tree. 
An example of this 
procedure for a simple noise process is shown in \Fref{f:syndrome}.
%and the corresponding braid moves are shown in \Fref{f:syndrome}(e).
The combinatorial derivation of the algorithm and data structures used for 
this procedure will be discussed elsewhere.
%Braiding processes that encircle 
%a non-trivial loop of our (toric) manifold can also be 
%treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
\cggb{what does the combinatorial comment mean?}
\cggb{why did the Pfeifer ref get commented? it should certainly go somewhere.}
%\cggb{This paragraph and the next still needs another look.}

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}
\caption{
(a) Noise processes initially form isolated sets of pair-created anyons, 
each crossing the boundary of a tile. 
(b) To measure the total charge 
contained within each tile, 
we first join the participating curve 
diagrams arbitrarily into a single curve diagram.
(c) The total charge of each tile can then be found 
by braiding anyons around each other until all charges within 
a tile are neighbors on the curve, as in (d).  
The red lines correspond to the worldlines for these braids.
}
\label{f:syndrome}
\end{center}
\vspace{-10pt}
\end{figure}

Following the results of the charge measurements, the decoder determines 
a recovery operation that involves fusion of subsets of anyons. 
These fusions can be simulated and their results calculated in the same way, 
and the output charge placed at an appropriate point in the lattice. 
This procedure is iterated until either the decoder terminates successfully, 
or the simulation itself declares failure.

@@

We plot the performance of the decoder as a function of noise strength for varying lattice sizes in 
\Fref{f:threshold}. 
The noise strength is parameterized by the Poisson process duration $t_{\mathrm{sim}}$, representing the expected number of errors per edge during the simulation. 
We find evidence of a decoding threshold below which decoding succeeds with asymptotic 
certainty as the system size increases at $t_{\mathrm{sim}}\simeq 0.125 \pm 0.003$.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated by an optimal decoder. 
However, calculations of such nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed result in a logical or leakage error.
As such it is likely that neglecting the full simulation of 
such events does not significantly affect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon model. 
Though we have chosen several properties of our model and 
simulation in a convenient way for simplicity,
Ref.~\cite{Brell2013} presents good evidence that it is
unlikely these choices will affect our results qualitatively.
%it seems highly
%unlikely that they will affect the qualitative results, similarly to the results of Ref.~\cite{Brell2013}. 
In particular, although we have modeled our logical qubit as 
encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred anyons. 
This situation would be appropriate to model error-correction routines for topological quantum computation. 
Additionally, we expect our results to be stable to changes in details of the noise model and decoding algorithm.

None of our techniques are restricted to simulation of Fibonacci anyon dynamics, and could 
equally well be used to simulate successful error-correction protocols in an arbitrary anyon code. 
As such, our methods could be used to demonstrate successful 
error-correction for an arbitrary anyonic topological quantum computer.

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, 
applying similar methods to more realistic models such as concrete 
microscopic spin models or models with non-topological features would give 
direct insight into practical error rates needed for topological quantum memories in nonabelian systems. 
It would also be interesting to find \emph{classical} spin models 
whose phase diagram encodes the threshold for error correction in these systems~\cite{Dennis2002}.
%Another potential direction is to incorporate tensor network methods for 
%simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
%Finally, perhaps the most important open question is whether 
Finally, an important extension of this work would be to the simulation of \emph{fault-tolerant} error-correction protocols for nonabelian anyon codes.
